---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)
```

# Beer Advocate Data Analysis

<!-- badges: start -->
<!-- badges: end -->

### Note on `rust-raven` project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [**codename**](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change. The seed `714062` was generated with random atmospheric noise at [random.org](https://www.random.org/).

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.4.0
codename(seed = 714062, type = "ubuntu")
#> [1] "rust raven"
```


## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [GNU Affero GPL](LICENSE).

## Credit

This README was adapted from what [Andrew Heiss](https://github.com/andrewheiss)
uses for a lot of his README files.